Research Spike: Evaluation of Using a Single Mnemonic both for Cryptocurrency and Identity Wallets

### 2021-12-13 prism-backend/research/single-mnemonic/discussion.tex
Initial content on the single mnemonic research spike. Summarizing BIP32.

\comment{This section is just a placeholder for now...}

Interesting topic for discussion, which applies mostly for ACs, but probably
to most other equivalent tools: we have tools that provide a (reasonably
complete and) secure way to authenticate in a privacy-preserving manner.
However, what do we do after such authentication takes place? E.g., it is of
little use to employ an AC to log into an online service, if that service
will learn all about us after we have anonymously authenticated. How does
privacy-preserving authentication improve the situation even in that case?
Is there some way in which we can help avoid that?

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End:

>>>>

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End:

### 2021-12-13 prism-backend/research/single-mnemonic/introduction.tex
Initial content on the single mnemonic research spike. Summarizing BIP32.

Privacy is a human right \needcite. It is also one of the most threatened ones
as a consequence of digitalization and hyperconnectivity \needcite. Maybe one
of the reasons behind this is that, on many occasions, it conflicts with another
human right: security \needcite. The most representative occasion in which this
happens is when we, as individuals%
\footnote{This, of course, also happens when authenticating machines -- as they
  carry to some extent information of some individual(s).}%
, want to gain access to some (possibly
digital) good or service. In that context, one has to prove having the legitimate
rights to access that good or service. However, in the process of doing so, the
entity verifying that we have the rights, typically learns something about us,
like our real name. This has been traditionally fine in physical systems, as the
impact was mostly local. However, with the advent of digitalization and
hyperconnectivity, this usually means that something related to our identity is
being logged in some server, who knows where and with what security controls,
and probably vulnerable to a number of criminal activities \needcite.

Of course, the cryptographic community has not been stranger to the need of
providing satisfactory solutions that enable both privacy and security. We
refer to all works trying to address to this need as \emph{privacy-preserving
  authentication} mechanisms. For decades, a huge amount of research, with many
variations, has been produced in the field. Sadly, still with limited impact in
the real world. Within this context, the goal of this review is manyfold: from a
practical perspective, serve
as a hub to security and privacy professionals that may find themselves in need
to balance privacy and security in some real world application; from a more
academic one, provide a (somehow detailed) overview of the vast amount of
literature in the field, which is already overwhelming enough to make it hard
to find the most appropriate variant of some property we may find ourselves
in need to provide; and finally, attempt to throw some insight into why such a
vast academic work has so far failed to gain more real world adoption.

In order to achieve this, %
we go over the main cryptographic primitives used to build privacy-preserving
authentication mechanisms. For all of them, we first provide a high-level
intuition of the approach to protect users' privacy. Then, we overview the
most prominent security models since their inception, and complement this
with some relevant variations concerning the provided functionality. We also
summarise the main results from an efficiency standpoint, and the main
challenges found so far both in theory and practice. To gain some initial
insight into the academic impact of each primitive, we provide (approximate)
numbers on publications in the main cryptography, security and privacy venues.
Finally, with a more practical mindset, we overview the applications that each
primitive has ``in theory'', and the extent to which they have been deployed
in the real world.

\note{Argue/mention that focus is on primitives that can be used for signing.
  This therefore excludes protocols that can only be used for anonymous
  authentication, but not anonymous singing per se (e.g., PAKE, PASTA, PESTO...
  and probably also Nymble, PEREA, BLAC...)}

\note{Create a graph of relationships between basic primitives and their
  applications. E.g., much of
  the work in SRP group signatures and DAA seems to be derived from randomizable
  signatures (which, afaik, are also a foundation for anonymous credentials).}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "sok-privsig"
%%% End:

>>>>

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "single-mnemonic"
%%% End:

### 2021-12-13 prism-backend/research/single-mnemonic/single-mnemonic.tex
Initial content on the single mnemonic research spike. Summarizing BIP32.

\documentclass{llncs}%[10pt]{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
\usepackage{amssymb}
\usepackage{xspace}
\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{State of the Art Review on Privacy-Preserving Authentication}
\author{}

@@

\begin{abstract}
  \comment{\textbf{DISCLAIMER}: Please, keep in mind that this is a working
    draft. Its main use in its current state is to foster debate on the topics
    it deals with. Do not take anything written here as ground truth.}

  This document explores three closely related cryptographic systems in the area
  of privacy-enhanced and, more concretely, in the domain of authentication.
  Namely, we study advances in group signatures (GS), anonymous credentials
  (AC), and direct anonymous attestation (DAA). All three have somehow similar
  purpose, have followed similar approaches, and have received much attention in
  the the academic literature for the last (almost) four decades but, in some
  sense, independently on one another.

  Our goal is to extract common insight from all of them. Firstly, from a
  theoretical perspective, with the aim to identify the advantages and
  disadvantages of existing models and approaches, and develop intuition as to
  whether we can (or cannot) combine ideas and fill gaps, and what may be the
  most suitable path to follow for new projects, depending on their needs.
  Secondly, from a practical perspective, DAA and ACs have already been deployed
  in the real world, and analysing the difficulties they have come across will
  certainly be of use.

\end{abstract}

\iffalse
\input{introduction.tex}
\fi
\input{gs.tex}
%\input{abs.tex}
\input{ac.tex}
\input{daa.tex}
\input{other.tex}
\input{discussion.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{soa-review}

\appendix

\section{TODOs}

General/pending doubts and tasks:

\begin{itemize}
\item Many constructions of ACs and GSs seem to be built using the same
  approach: an issuer and a user interact so that the latter receives a signature
  from the former; then the user employs that signature to do something else (in
  ACs, prove attributes; in GSs, produce further signatures). Still, the typical
  security models are different. Concretely, GSs have a very well established
  model, but ACs don't. It also ``bothers'' me that none of the AC models I've
  seen cover something similar to non-frameability, nor traceability, but just
  include conventional unforgeability properties. Why is this?
  \begin{itemize}
  \item Concerning non-frameability, this probably is because the de-anonymizing
    functionality that ACs include (if any) is revocation, which is essentially
    a blacklisting. On the other hand, GSs usually allow opening, which extracts
    the actual identity. Opening is ``stronger'' than blacklisting, in the sense
    that it can be used to build blacklisting, but it is also more threatening
    to privacy and may have further implications. Among these, unequivocable
    framing of users. This, in general, does not seem possible with conventional
    blacklisting (or, at least, not in an unequivocable way).
  \item Concerning traceability. \cite{fhs19} definition of unforgeability
    resembles traceability. I guess this would be the case in most AC works.
  \end{itemize}
\end{itemize}

\input{app-gs}
\input{app-ac}

>>>>

\documentclass{llncs}
\pagestyle{plain}

\usepackage[n,advantage,operators,sets,adversary,landau,probability,notions,
logic,ff,mm,primitives,events,complexity,asymptotics,keys]{cryptocode}
%\usepackage{amssymb}
\usepackage{xspace}
%\usepackage[normalem]{ulem}
\usepackage{hyperref}

\title{Research Spike: Evaluation of Using a Single Mnemonic both
  for Cryptocurrency and Identity Wallets}
\author{}

@@

\begin{abstract}
  \comment{\textbf{DISCLAIMER}: Please, keep in mind that this is a working
    draft. Its main use in its current state is to foster debate on the topics
    it deals with. Do not take anything written here as ground truth.}  
\end{abstract}

\input{introduction.tex}
\input{bip32-etal.tex}
\input{idwal.tex}
\input{discussion.tex}
\input{conclusion.tex}

\bibliographystyle{splncs04}
\bibliography{single-mnemonic}
